---
title: "An Analysis of BC Grade-to-Grade Transitions"
author: "Warren Zhu"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up everything here. This is hidden from the final output
#install.packages(c("data.table","leaflet"))
library(readxl)
library(data.table)
library(leaflet)
library(tidyverse)
library(ggplot2)
library(mgcv)

library(plotly)
library(knitr)
library(widgetframe)
```

## Introduction

This site was created with the intention of answering the following question: 

**How has the quality of education/opportunities in the past 30 years differed in different kinds of schools, as well as for different types of British Columbian students (i.e., indigenous? Special needs?)?**

Note that although this is just one question, it is a very complex one, and so naturally, a lot of data is required to answer it. Thus, this website uses [open data from British Columbia](https://catalogue.data.gov.bc.ca/)--specifically, [the 1992-2021 data on grade-to-grade transition published by British Columbia Education Analytics](https://catalogue.data.gov.bc.ca/dataset/bc-schools-grade-to-grade-transition). To gather information regarding the districts, this site also uses [information regarding each district's office, also published by the British Columbia Education Analytics](https://catalogue.data.gov.bc.ca/dataset/bc-schools-school-district-information). By using the grade-to-grade transitions as a metric for the quality of education/opportunities, we can formulate a nuanced response to the question.

If you wish to see the Github repository for this site, [you can find it here](https://github.com/CoderWarren/An-Analysis-of-BC-Grade-Transitions).

$$\\[0.2in]$$

## Methods

### Data Prep

```{r read-in, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the 1992 to 2001 data
data_1 <- read_excel("data_transitions/grtogrtransition_1992-93_to_2000-01.xlsx")
# Read in the 2001 to 2011 data
data_2 <- read_excel("data_transitions/grtogrtransition_2001-02_to_2010-11.xlsx")
# Read in the 2011 to 2021 data
data_3 <- read_excel("data_transitions/grtogrtransition_2011-12_to_2020-21.xlsx")
# Read in the locations
locations <- data.table::fread("data_districts/current_district_office_information.csv")
```

```{r data-prep-1, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
# Here, we prepare the merged data set

# Combine all the data
data <- rbind(data_1, data_2, data_3)

# Remove all missing values
data <- data[!(data$DENOMINATOR_STUDIED_COHORT=="Msk"|
                 data$NUMERATOR_NUM_SUCCESSFUL=="Msk"|
               data$G2G_TRANSITION_RATE=="Msk"),]

# Make an integer variable storing the year
data$year <- strtoi(substr(data$SCHOOL_YEAR, 1, 4))

# Make the grade to grade transition rates a float
data$rate <- as.double(data$G2G_TRANSITION_RATE)

# Make the number of students an integer
data$num_students <- as.integer(data$DENOMINATOR_STUDIED_COHORT)

# Make the number of students who successfully moved on an integer
data$num_successful <- as.integer(data$NUMERATOR_NUM_SUCCESSFUL)

# Make the student grade a factor
data$grade <- as.factor(data$STUDENT_GRADE)

# Drop the unnecessary columns
data <- subset(data, select=-c(BASE_YEAR, SCHOOL_YEAR, G2G_TRANSITION_RATE,
                               MEASURE_NAME, DENOMINATOR_STUDIED_COHORT,
                               NUMERATOR_NUM_SUCCESSFUL, STUDENT_GRADE,
                               DISTRICT_NAME))
```

```{r data-prep-2, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
data %>% 
  filter(DATA_LEVEL == "Province Level") %>%
  filter(FACILITY_TYPE == "All Facility Types") -> data_province
```

```{r data-prep-3, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
# Drop all the rows in our table regarding data on a "Province Level"
# This is because we will not use it for the remainder of the assignment
data_district <- data %>% filter(DATA_LEVEL == "District Level")
data_district$district <- as.integer(data_district$DISTRICT_NUMBER)

# Now that we're on a district level, we can use the location data
data_district <- left_join(x=data_district, y=locations, 
                           by=c("district" = "DISTRICT_NUMBER"))

# Drop columns that aren't used
data_district <- subset(data_district, select=-c(
  DISTRICT_PROVINCE, DISTRICT_POSTAL_CODE, 
  DISTRICT_PHONE_NUMBER, DISTRICT_WEBSITE))

# Turn our district into a factor variable
data_district$district <- as.factor(data_district$district)

# Check what we have
summary(data_district)
```

```{r data-prep-4, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
data_school <- data %>% filter(DATA_LEVEL == "School Level")
data_school$district <- as.integer(data_school$DISTRICT_NUMBER)
data_school$year <- as.factor(data_school$year)
```

Our data came in the form of 3 '.xlsx' files containing information regarding BC grade-to-grade transitions on a provincial, district and school level, while the data regarding the location of the offices of each district was in the form of a '.csv' file. The three '.xlsx' files each contained information on different periods: one file contained information from 1992/93 to 2000/01, another contained information from 2000/01 to 2010/11, and our last one contained information from 2010/11 to 2020/21. There are many rows in each of the '.xlsx' files (the first had 374654, the second had 541501, and the third had 536978), but it is also worth noting that the data was not tidy, as some rows recorded data on "all students" whilst other rows recorded data on "Indigenous students", so each row should not be treated as its own observation. Then, we combined our data into one large data table, and removed entries in our table with missing data.

### Procedure

Before doing anything meaningful, we should first do some exploratory data analysis. Based on the structure of the data, I felt as though it would be most appropriate to look at the situation from all three levels. We begin by making interactive figures to understand the situation from a provincial and district level. We will also perform model fitting on a school level. By mimicking the structure of our data, we can make the most out of it!

```{r}
#WORK IN PROGRESS (A lot of the existing work is from the Midterm)
```

$$\\[0.2in]$$

## Results

Prior to doing anything, let's do some exploratory data analysis. You can see the plots and analysis that were made for the provincial, district and school levels in the corresponding tabs.

### Exploratory Data Analysis {.tabset}

#### Provincial Level

Below is an interactive plot. The axis is fixed for ease of comparison, but you can always zoom in. We can observe various things by tweaking the settings of the plot. Below are just a few discoveries that I found.

* In the `Province-Total` and `All Students` setting, we can see that overall, as time went by, the percentage of students that successfully transitioned to the next grade increased. This is the most prominent for students in grades 8-11. It is also worth noting that the percentage of students in grades 10/11 that successfully transition to the next grade are significant lower than the percentage of students in the lower grades that successfully transition.
* Keeping the bottom dropdown as `All Students` but altering the top dropdown between `BC Public School` and `BC Independent School`, we can see that the rates of grade transitions for students of BC public schools are lower than that of BC independent schools, but the increase in the rate of successful grade transitions is significantly higher in public schools. Furthermore, private schools (i.e., independent schools) seem to have a much higher rate of transition, especially for grade levels 8 and above. This does make sense, as private schools are often times more expensive, and so the students attending them naturally live in my affluent families.
* Keeping the top dropdown as `Province-Total` but altering the bottom dropdown between `Indigenous` and `Non Indigenous`, we can see that the rate of successful grade transitions is significantly lower in Indigenous students compared to their non-Indigenous counterparts. We can also see that the increase in the rates of grade transitions over the years is higher for Indigenous students compared to non-Indigenous students.
* Keeping the top dropdown as `Province-Total` but altering the bottom dropdown between `Non Special Needs` and `Special Needs`, we can see that the rate of successful grade transitions is significantly lower in special needs students compared to students without special needs.

$$\\[0.1in]$$

```{r plot-prep1, echo=FALSE, message=FALSE, warning=FALSE}
province_plot <- data_province %>%  # Code for the plot
  plot_ly(x = ~year, 
          y=~rate, 
          type="scatter", 
          mode="line",
          customdata = ~SUB_POPULATION,
          ids = ~PUBLIC_OR_INDEPENDENT,
          color=~grade, 
          colors="Set2",
          hoverinfo = 'text',
          text = ~paste('</br> Year: ', year,
                        '</br> Grade: ', grade,
                      '</br> Rate: ', rate),
          transforms=list(
            list(
              type = 'filter',
              target = 'customdata',
              operation = '=',
              value = unique(data_province$SUB_POPULATION)[1]
              ),
            list(
              type = 'filter',
              target = 'ids',
              operation = '=',
              value = unique(data_province$PUBLIC_OR_INDEPENDENT)[1]
              )
            )
          ) %>% # The dropdown menu for SUBPOPULATION
  layout(
    title="Rate of successful grade transitions (1993-2020)",
    yaxis = list(title = "% of Students who Transitioned to the Next Grade",
                 range = c(40,100)),
    xaxis=list(title="Year",
               categoryorder = "array",
              categoryarray = data_province$year),
    legend = list(title=list(text='<b> Student Grade Level </b>')),
    updatemenus = list(
      list(
        x=1,
        y = 0.1,
        type = 'dropdown',
        direction="up",
        active = 0,
        buttons=list(
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_province$SUB_POPULATION)[1]),
               label = unique(data_province$SUB_POPULATION)[1]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_province$SUB_POPULATION)[2]),
               label = unique(data_province$SUB_POPULATION)[2]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_province$SUB_POPULATION)[3]),
               label = unique(data_province$SUB_POPULATION)[3]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_province$SUB_POPULATION)[4]),
               label = unique(data_province$SUB_POPULATION)[4]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_province$SUB_POPULATION)[5]),
               label = unique(data_province$SUB_POPULATION)[5])
          )
        ),
      list(
        x=1,
        y = 0.2,
        type = 'dropdown',
        direction="up",
        active = 0,
        buttons=list(
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_province$PUBLIC_OR_INDEPENDENT)[1]),
               label = unique(data_province$PUBLIC_OR_INDEPENDENT)[1]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_province$PUBLIC_OR_INDEPENDENT)[2]),
               label = unique(data_province$PUBLIC_OR_INDEPENDENT)[2]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_province$PUBLIC_OR_INDEPENDENT)[3]),
               label = unique(data_province$PUBLIC_OR_INDEPENDENT)[3])
          )
        )
  ))

# Had a lot of trouble creating this plot, but this link helped:
# https://stackoverflow.com/questions/52321695/filter-a-plotly-line-chart-based-on-categorical-variable
```

```{r plot1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 1: An interactive graph depicting the percentage of students in BC of each grade level that successfully transition to the next grade between the years 1993 and 2020."}
province_plot
```

$$\\[0.1in]$$

#### District Level

Below are two interactive plots. Figure 2 displays the individual data for each of the districts, whilst Figure 3 consists of the summary of all the districts for each year. The dropdown menus for both plots allow you to select a specific grade and population. The axis is fixed for ease of comparison, but you can always zoom in (or hover your mouse to see more details). There are many interesting things you can find with the interactive plots. Below are just a few findings:

* The percentage of students who transitioned to the next grade for Grade 11 fluctuates very dramatically. Playing with the interactive boxplot, one can see that the variation in the rate of transitions from grade 11 to 12 from 2015 to 2020 has been significantly reduced (when compared with the past years). The median of the rates has also increased over time. As such, these plots suggest that there is more equality (due to the decrease in variation) in terms of education/opportunities across the districts, and the education/opportunities offered is of a higher quality (as indicated by the higher median).

$$\\[0.1in]$$

```{r plot-prep2, echo=FALSE, message=FALSE, warning=FALSE}
district_plot_1 <- data_district %>%  # Code for the plot
  plot_ly(x = ~year, 
          y=~rate, 
          type="scatter", 
          mode="line",
          customdata = ~SUB_POPULATION,
          ids = ~grade,
          color=~district, 
          colors="Set2",
          hoverinfo = 'text',
          text = ~paste('</br> Year: ', year,
                        '</br> District: ', district,
                        '</br> Grade: ', grade,
                      '</br> Rate: ', rate),
          transforms=list(
            list(
              type = 'filter',
              target = 'customdata',
              operation = '=',
              value = unique(data_district$SUB_POPULATION)[1]
              ),
            list(
              type = 'filter',
              target = 'ids',
              operation = '=',
              value = unique(data_district$grade)[1]
              )
            )
          ) %>% # The dropdown menu for SUBPOPULATION
  layout(
    title="Rate of successful grade transitions (1993-2020)",
    yaxis = list(title = "% of Students who Transitioned to the Next Grade",
                 range = c(20,100)),
    xaxis=list(title="Year",
               categoryorder = "array",
              categoryarray = data_province$year),
    legend = list(title=list(text='<b> District </b>')),
    annotations = list(text = "<b> Grade: </b>",  x = 0.9, xref="paper",
                       y = 0.13, yref="paper", showarrow=FALSE ),
    updatemenus = list(
      list(
        x=1,
        y = 0.1,
        type = 'dropdown',
        direction="up",
        active = 0,
        buttons=list(
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_district$SUB_POPULATION)[1]),
               label = unique(data_district$SUB_POPULATION)[1]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_district$SUB_POPULATION)[2]),
               label = unique(data_district$SUB_POPULATION)[2]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_district$SUB_POPULATION)[3]),
               label = unique(data_district$SUB_POPULATION)[3]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_district$SUB_POPULATION)[4]),
               label = unique(data_district$SUB_POPULATION)[4]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_district$SUB_POPULATION)[5]),
               label = unique(data_district$SUB_POPULATION)[5])
          )
        ),
      list(
        x=1,
        y = 0.2,
        type = 'dropdown',
        direction="up",
        active = 0,
        buttons=list(
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[1]),
               label = unique(data_district$grade)[1]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[2]),
               label = unique(data_district$grade)[2]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[3]),
               label = unique(data_district$grade)[3]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[4]),
               label = unique(data_district$grade)[4]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[5]),
               label = unique(data_district$grade)[5]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[6]),
               label = unique(data_district$grade)[6]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[7]),
               label = unique(data_district$grade)[7]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[8]),
               label = unique(data_district$grade)[8]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[9]),
               label = unique(data_district$grade)[9]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[10]),
               label = unique(data_district$grade)[10]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[11]),
               label = unique(data_district$grade)[11])
          )
        )
  ))
```

```{r plot2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2: An interactive graph depicting the percentage of high school BC students that successfully transition to the next grade between the years 1993 and 2020."}
district_plot_1
```

$$\\[0.1in]$$

```{r plot-prep3, echo=FALSE, message=FALSE, warning=FALSE}
data_district$year <- as.factor(data_district$year)
district_plot_2 <- data_district %>%
  plot_ly(
    y=~rate,
    color=~year,
    type="box",
    customdata=~SUB_POPULATION,
    ids=~grade,
    transforms=list(
      list(
        type = 'filter',
        target = 'customdata',
        operation = '=',
        value = unique(data_district$SUB_POPULATION)[1]
        ),
      list(
        type = 'filter',
        target = 'ids',
        operation = '=',
        value = unique(data_district$grade)[1]
        )
      )
    ) %>% # The dropdown menu for SUBPOPULATION
  layout(
    title="Rate of successful grade transitions (1993-2020)",
    yaxis = list(title = "% of Students who Transitioned to the Next Grade",
                 range = c(20,100)),
    xaxis=list(title="Year",
               categoryorder = "array",
              categoryarray = data_province$year),
    annotations = list(text = "<b> Grade: </b>",  x = 0.9, xref="paper",
                       y = 0.13, yref="paper", showarrow=FALSE ),
    legend = list(title=list(text='<b> Year </b>')),
    updatemenus = list(
      list(
        x=1,
        y = 0.1,
        type = 'dropdown',
        direction="up",
        active = 0,
        buttons=list(
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_district$SUB_POPULATION)[1]),
               label = unique(data_district$SUB_POPULATION)[1]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_district$SUB_POPULATION)[2]),
               label = unique(data_district$SUB_POPULATION)[2]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_district$SUB_POPULATION)[3]),
               label = unique(data_district$SUB_POPULATION)[3]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_district$SUB_POPULATION)[4]),
               label = unique(data_district$SUB_POPULATION)[4]),
          list(method = "restyle",
               args = list("transforms[0].value", unique(data_district$SUB_POPULATION)[5]),
               label = unique(data_district$SUB_POPULATION)[5])
          )
        ),
      list(
        x=1,
        y = 0.2,
        type = 'dropdown',
        direction="up",
        active = 0,
        buttons=list(
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[1]),
               label = unique(data_district$grade)[1]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[2]),
               label = unique(data_district$grade)[2]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[3]),
               label = unique(data_district$grade)[3]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[4]),
               label = unique(data_district$grade)[4]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[5]),
               label = unique(data_district$grade)[5]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[6]),
               label = unique(data_district$grade)[6]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[7]),
               label = unique(data_district$grade)[7]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[8]),
               label = unique(data_district$grade)[8]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[9]),
               label = unique(data_district$grade)[9]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[10]),
               label = unique(data_district$grade)[10]),
          list(method = "restyle",
               args = list("transforms[1].value", unique(data_district$grade)[11]),
               label = unique(data_district$grade)[11])
          )
        )
  ))
```

```{r plot3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3: An interactive boxplot depicting the percentage of high school BC students from each district that successfully transitioned to the next grade between the years 1993 and 2020."}
district_plot_2
```

$$\\[0.1in]$$

#### School Level

```{r}
#WORK IN PROGRESS
```

$$\\[0.1in]$$

## Conclusions and Summary

```{r}
#WORK IN PROGRESS
```
