---
title: "An Analysis of BC Grade-to-Grade Transitions"
author: "Warren Zhu"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up everything here. This is hidden from the final output
#install.packages(c("data.table","leaflet"))
library(readxl)
library(data.table)
library(leaflet)
library(tidyverse)
library(ggplot2)
library(mgcv)
```

```{r read-in, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the 1992 to 2001 data
data_1 <- read_excel("data_transitions/grtogrtransition_1992-93_to_2000-01.xlsx")
# Read in the 2001 to 2011 data
data_2 <- read_excel("data_transitions/grtogrtransition_2001-02_to_2010-11.xlsx")
# Read in the 2011 to 2021 data
data_3 <- read_excel("data_transitions/grtogrtransition_2011-12_to_2020-21.xlsx")
# Read in the locations
locations <- data.table::fread("data_districts/current_district_office_information.csv")
```

## Introduction

This site was created with the intention of answering the following question: 

**How has the quality of education/opportunities in the past 30 years differed in different kinds of schools, as well as for different types of British Columbian students (i.e., indigenous? Special needs?)?**

Note that although this is just one question, it is a very complex one, and so naturally, a lot of data is required to answer it. Thus, this website uses [open data from British Columbia](https://catalogue.data.gov.bc.ca/)--specifically, [the 1992-2021 data on grade-to-grade transition published by British Columbia Education Analytics](https://catalogue.data.gov.bc.ca/dataset/bc-schools-grade-to-grade-transition). To gather information regarding the districts, this site also uses [information regarding each district's office, also published by the British Columbia Education Analytics](https://catalogue.data.gov.bc.ca/dataset/bc-schools-school-district-information). By using the grade-to-grade transitions as a metric for the quality of education/opportunities, we can formulate a nuanced response to the question.

If you wish to see the Github repository for this site, [you can find it here](https://github.com/CoderWarren/An-Analysis-of-BC-Grade-Transitions).

## Methods



## Results

```{r}
#WORK IN PROGRESS
```

## Conclusions and Summary

```{r}
#WORK IN PROGRESS
```
